Hydras \& Co.

### 2022-02-28 doc/gaia-chapter.tex
gaia-hydras first draft

\chapter{Gaia and the hydras (experimental)}
\index{coq}{Plug-ins!Gaia}
\label{gaia-chapter}

@@

@@

Hence, both definitions of ``being in Cantor normal form'' are
compatible.

@@

@@

\section{Importing theorems from \HydrasLib}

Many lemmas of the library \texttt{ordinals/Epsilon0/} are statements about inhabitants of type \texttt{hT1}. In order to get
similar statements on \gaia's type \texttt{T1}, we prove and
apply rewriting rules of the following forms:

\vspace{4pt}

\noindent\emph{From Module~\href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge}}

\inputsnippets{T1Bridge/T1leIff}
\inputsnippets{T1Bridge/rewritingRules}
\inputsnippets{T1Bridge/T1ltIff}

@@

These tools allow us to import \texttt{E0.lt}'s  well-foundedness.

@@

\subsection{E0 as an ordinal notation}
The module \href{../theories/html/gaia.T1Bridge.html}{gaia.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
(see Chapter~\ref{chap:ON}).

@@

@@

The library \href{../theories/html/gaia.GPaths.html}%
{\texttt{gaia.GPaths}} imports definitions and lemmas from
\href{../theories/html/hydras.Epsilon0.Paths.html}%
{\texttt{hydras.Epsilon0.Paths}}.

\subsubsection{Remark}
\label{sect:gaia-masking}

First, we import some definitions. When we import some constant \emph{Id} from {\texttt{hydras.Epsilon0.Paths}} the name
\emph{Id} is given to {\texttt{gaia.GPaths}}. In order to get
{\texttt{hydras.Epsilon0.Paths}}'s sense, please refer it as
\emph{Paths.Id}.
Sometimes, we rename \emph{Paths.Id} into \emph{{\bf h}Id}.

>>>>

\chapter{Gaia and the hydras (draft)}
\index{coq}{Plug-ins!Gaia}
\label{gaia-chapter}

@@

Refinement lemmas can be easily ``reversed''.

@@

\subsubsection{More rewriting lemmas}

\inputsnippets{T1Bridge/T1ltIff}
\inputsnippets{T1Bridge/T1leIff}

\inputsnippets{T1Bridge/rewritingRules}

@@

Both definitions of ``being in Cantor normal form'' are
compatible.

@@

@@

These tools allow us, for instance,  to import \texttt{E0.lt}'s  well-foundedness.

@@

\section{E0 as an ordinal notation}
The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
(see Chapter~\ref{chap:ON}).

@@

\section{Importing Definitions and theorems from \HydrasLib}

Some constructions of \HydrasLib were (to our knowledge) not implemented in \gaia yet. For instance, commutative (``Hessenberg'') addition, canonical sequences and Ketonen-Solovay machinery (Chapters~\ref{chap:T1}, \ref{chap:ketonen} and \ref{chap:alpha-large}). In order not to copy long proofs, we chose to
\emph{derive} several constructions from \HydrasLib into \gaia's world, and rewrite \HydrasLib statements in order to make them \gaia-compatible.

\subsubsection{Remark}
\label{sect:gaia-masking}
We try to respect the following ``protocol''.
Let us assume we adapt some definitions and properties
from a given module \texttt{hydras.Epsilon0.{\it XXX}}.
\begin{enumerate}
\item We create a module \texttt{gaia\_hydras.{\it GXXX}},
\item When we import some constant \emph{Id} from
 \texttt{hydras.Epsilon0.{\it XXX}}, the name 
\emph{Id} is given to \texttt{gaia\_hydras.{\it GXXX.Id}}.
In order to get
\texttt{hydras.Epsilon0.{\it XXX}}'s sense, please refer it as
  \emph{XXX.Id} or \emph{Epsilon0.XXX.Id}.
  \item Sometimes, we rename \emph{XXX.Id} into \emph{{\bf h}Id} (for readability's sake).
\end{enumerate}

\subsection{Hessenberg sum}

Natural sum  (a.k.a. Hessenberg's  sum) is defined in
\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).

We define this function by delegation to \HydrasLib'.
\vspace{4pt} 
\noindent.

From\href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg}

\inputsnippets{GHessenberg/oplusDef}

By simple rewriting, we import the following lemmas.

\inputsnippets{GHessenberg/oplusEquations}

Hessenberg sum is associative, commutative and strictly monotonous (on ordinal terms in normal form).

\inputsnippets{GHessenberg/oplusLemmasa}
\inputsnippets{GHessenberg/oplusLemmasb}
\inputsnippets{GHessenberg/oplusLemmasc}
\inputsnippets{GHessenberg/oplusLemmasd}
\inputsnippets{GHessenberg/oplusLemmase}

@@

The library \href{../theories/html/gaia_hydras.GPaths.html}%
{\texttt{gaia\_hydras.GPaths}} imports definitions and lemmas from
\href{../theories/html/hydras.Epsilon0.Paths.html}%
{\texttt{hydras.Epsilon0.Paths}}.
